---
title: 'Chick Growth by Diet '
author: "Mark Pratley"
date: "1 December 2015"
output: 
  slidy_presentation: 
    self_contained: no
---

```{r lib, echo=F, warning=F, message=FALSE}
library(ggvis)
library(shiny)
library(dplyr)
library(ggplot2)

# our dataset
cw <- ChickWeight
```

## Chick Weight
### A (Non-interactive) Shiny Presentation

This is a practise presentation using shiny, ggplot and ggvis to explore the ChickWeight dataset.

To see this as a presentation click [here](../ChickWeight-Presentation.html)

The ChickWeight dataset has 578 rows and 4 columns from an experiment on the effect of diet on early growth of chicks.

First we will explore the dataset and then we will compare the effect of diet on chick weight.

## Variables

* Weight
    + A numeric vector giving the body weight of the chick (gm).

* Time
    + A numeric vector giving the number of days since birth when the measurement was made.

* Chick
    + An ordered factor with levels 18 < ... < 48 giving a unique identifier for the chick.

* Diet
    + A factor with levels 1-4 indicating which experimental diet the chick received.

## Summary of the Data

```{r data}
summary(cw)
```

## Weight over Time (Hoverable)

```{r weight_t, echo=F}
all_values <- function(x) {
  if(is.null(x)) return(NULL)
  paste0(names(x), ": ", format(x), collapse = "<br />")
}
cw %>% 
  ggvis(x=~Time, y=~weight) %>% 
  add_tooltip(all_values, "hover") %>% 
  layer_boxplots()
```

## Weight Density Plot

```{r wgt_den, echo=F}
cw %>% ggvis(x = ~weight) %>% layer_densities()
```

## Growth Plot for Individual Chicks

```{r growth, echo=FALSE}
cw %>% ggplot(aes(x=Time, y=weight, colour=Diet, group=Chick)) +
        geom_line()
```

## Summary

From our quick data exploration we can see that there are:

* 50 chicks
* On 4 Diets
* Over 21 days
  
* Weight
    + Range 0 < w < 400
    + Mean of 121
    + Median of 103
    + Standard deviation of 71

## Scatter Plot of Growth Data

```{r growth-scatter, echo=FALSE}
all_values <- function(x) {
  if(is.null(x)) return(NULL)
  paste0(names(x), ": ", format(x), collapse = "<br />")
}
cw %>% ggvis(x = ~Time, y = ~weight) %>%
  layer_points(fill.hover := "red") %>%
  add_tooltip(all_values, "hover") %>% 
  layer_smooths()

```

## Scatter Plot of Data Grouped by Diet

```{r growth_scatter_diet, echo=F, warning=F, message=FALSE}
cw %>% ggplot(aes(x=Time, y=weight, colour=Diet)) +
        geom_point(alpha=.4) +
        geom_smooth(method = "loess", alpha=.2, size=1) +
        ggtitle("Chick Growth Grouped by Diet")
```

## BoxPlot of Weight Data Grouped by diet (Hoverable)

```{r weight_box, echo=F, warning=F, message=FALSE}
# Boxplot with hover values
cw %>% 
    ggvis(~Diet, ~weight, fill=~Diet ) %>% 
    layer_boxplots() %>% 
    add_tooltip(all_values, "hover") %>% 
    hide_legend("fill")
```

## Final Weight Density Grouped by Diet

```{r final_wgt_diet, echo=F, warning=F, message=FALSE}
cw %>% 
  filter(Time==21) %>% 
  ggplot(aes(x=weight, colour=Diet)) +
    geom_density()
```

## Final Weight Histogram Grouped by Diet

```{r final_wgt_hist, echo=F, warning=F, message=FALSE}
cw %>% 
  filter(Time>=21) %>% 
  ggplot(aes(x=weight, fill=Diet)) +
    geom_histogram(colour="black", binwidth=50) +
    facet_grid(Diet ~ .) +
    theme(legend.position="none")
```


## Survival Rate Grouped by Diet

```{r survival_diet, echo=F, warning=F, message=FALSE}
cw2 <- cw %>% 
        group_by(Time, Diet) %>% 
        summarise(Num=n(), Percent=n()/10)
cw2$Percent <- ifelse(cw2$Diet==1, cw2$Percent/2, cw2$Percent)

cw2 %>% 
  ggplot(aes(x=Time, y=Percent, fill=Diet)) +
    geom_bar(position="dodge", stat="identity")
```

## Summary of Final Weights
### Grouped by Diet

```{r final_wgt_summary, echo=F, warning=F, message=FALSE}
cw %>% 
    filter(Time>=21) %>%
    group_by(Diet) %>% 
    summarise(Num=n(), 
              min=min(weight), 
              max=max(weight), 
              mean=mean(weight),
              median=median(weight),
              sd=sd(weight)
              ) %>%
  as.data.frame() %>% 
  print()
```

## Conclusions

- Diet #3 gives the highest mean and median weight
- Diet #4 is the most consistent with the lowest standard deviation
- Diet #1 has the highest mortality rate
- Diet #4 has the biggest mean weight until day 12 - After which Diet #3 took over

In conclusion, for a fixed Diet, then #3 would be the best choice if selecting for mass, but if uniformity was the desired outcome, then Diet #4 would be the best choice.

It would be worth investigating whether a diet consisting of Diet #4 for the first 10-12 days followed by Diet #3 leads to an increased yield.